---
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "readme_code/",
  out.width = "100%"
)
```

<!-- badges: start -->
 [![R-CMD-check](https://github.com/klosins/dbc/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/klosins/dbc/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

# dbc

## Dynamic bias correction for panel data estimators

This package implements the dynamic biases correction estimator as per Klosin, S, Dynamic Biases of Static Panel Data Estimators (forthcoming). This allows the estimation of treatment effects in panel fixed effects models where the outcome is dynamic. See below for usage and examples.


## Installation

```r
# Install the latest version on github
pak::pak("github::klosins/dbc")
```

## Usage

```{r load}
library(dbc)
```

The follows shows example usage of the DBC under three general types of model. The variables of interest are the outcome variable, $Y_{it}$ for individual $i$ and time $t$, and the treatment variable $D_{it}$.

The first is the **exogenous** treatment model, where the outcome is:

$$Y_{it} = \alpha_{i} + \rho_{1} Y_{t-1} + \tau D_{t} + \beta_{1} X_{1} +\varepsilon_{it}$$

and with the treatment equation:

$$D_{it} = u_{it}$$

The following shows the simulation of this model using the `dbc::DGP` data generating process function, and then the estimation of this model.

`summary(fit_exog)` shows the bias-corrected estimates. The reported standard errors used underlying iid errors, and are not clustered.

The biased, OLS estimates of the model can be shown as `fit_exog$biased_coefficients`. `str(fit_exog)` shows the full list of objects returned with the `dbc` model.

```{r exog}

set.seed(42)

# Exogenous treatment (no lag_y in treatment equation)
data_exog <- DGP(N = 500, N_T = 4, rho1 = 0.2, rho2 = 0, tau = 0.5, n_X1 = 1, n_X2 = 0)

fit_exog <- dbc(
    outcome_fml = y ~ lag_y + D + X1_1,
    lag_y       = "lag_y",
    treatment   = "D",
    panel_id    = "panel_id",
    time_id     = "time",
    data        = data_exog
)


summary(fit_exog)
```

The second, is the **endogenous** treatment model, where the treatment effect is also a function of the previous values of the outcome:

$$Y_{it} = \alpha_{i} + \rho_{1} Y_{t-1} + \tau D_{t} \beta_{1}X_{1}  +\varepsilon_{it}$$

$$D_{it} = c_{i} + \rho_{2} Y_{t-1}  + \beta_{2} X_{2} + u_{it}$$


```{r endo}


# Endogenous treatment (lag_y in treatment equation)
data_endo <- DGP(N = 500, N_T = 4, rho1 = 0.2, rho2 = 0.3, tau = 0.5,n_X1 = 1, n_X2 = 1)

fit_endo <- dbc(
    outcome_fml   = y ~ lag_y + D,
    treatment_fml = D ~ lag_y,
    lag_y         = "lag_y",
    treatment     = "D",
    panel_id      = "panel_id",
    time_id       = "time",
    data          = data_endo
)

summary(fit_endo)
```

Last, the **interaction** model, which allows for interactions between the treatment variable and exogenous variables, $W$: 

$$Y_{it} = \alpha_{i} + \rho_{1} Y_{t-1} + \tau_{1} W_{1} \times D_{t} + \tau_{2} W_{2} \times D_{t} \beta_{1}X_{1}  +\varepsilon_{it}$$

$$D_{it} = c_{i} + \rho_{2} Y_{t-1}  + \beta_{2} X_{2} + u_{it}$$


```{r interact}


# With moderators/interactions
data_interact <- DGP(
    N = 500, N_T = 4,
    rho1 = 0.2, rho2 = 0.3,
    tau = c(0.8, 1.2),
    n_X1 = 1, n_X2 = 1, n_W = 2
)


fit_interact <- dbc(
    outcome_fml   = y ~ lag_y + D:W_1 + D:W_2 + X1_1,
    treatment_fml = D ~ lag_y + X2_1,
    lag_y         = "lag_y",
    treatment     = "D",
    panel_id      = "panel_id",
    time_id       = "time",
    data          = data_interact
)

summary(fit_interact)
```

### S3 methods


Standard model sumamry functions on the dbc output are provided.

Returning the bias-corrected coefficients:

```{r}
coef(fit_endo)
```


Returning the variance-convariance matrix on the bias-corrected coefficients:
```{r}
vcov(fit_endo)
```

Standard confidence intervals:
```{r}
confint(fit_endo, level = 0.90)
```

The number of observations:
```{r}
nobs(fit_endo)
```

### broom / modelsummary

The package has methods to interface with the `broom` and `modelsummary` packages.

Create a tidy version of the bias-corrected coefficients:

```{r broom}
library(broom)

tidy(fit_endo, conf.int = TRUE, conf.level = 0.95)

```

`glance` shows other model information, like number of observations.

```{r}
glance(fit_endo)
```

We can produce tables from `modelsummary`. Using the `shape = group + term ~ model` argument groups the table by whether the coefficients are in the outcome or treatment equation.

```{r modelsummary}
library(modelsummary)

# Multiple models side-by-side
modelsummary(
    list(Exogenous = fit_exog, Endogenous = fit_endo),
    shape = group + term ~ model
)
```
